\documentclass[10pt,twocolumn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{hyperref}

\title{A Galilean Frame-Tracking Method for TRML Simulations}
\author{AthenaK TRML Technical Note}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This document describes the frame-tracking method implemented in
\texttt{src/pgen/TRML.cpp} for TRML simulations. The method is a cadence-gated,
temperature-filtered, Galilean control loop that combines proportional,
derivative, and optional integral feedback. The controller is designed to keep
the interface near a user-defined target while preserving conservative
hydrodynamic updates, limiting transient over-correction, and maintaining robust
runtime behavior in turbulent flows.
\end{abstract}

\section{Purpose and design constraints}
TRML runs can develop long-time vertical drift of the thermally selected
interface. Left uncontrolled, drift can push the interface toward domain
boundaries and degrade physical interpretability. The implemented strategy
applies a \emph{uniform Galilean velocity shift} $\Delta v_z$ at discrete
actuation times. Two constraints drive the design:
\begin{enumerate}
  \item Control authority must act on a physically meaningful interface
  observable, not on instantaneous local fluctuations.
  \item The update must remain conservative in momentum and (for ideal EOS)
  total energy, and avoid impulsive forcing.
\end{enumerate}
The method therefore uses filtered global moments, optional bias cancellation,
and explicit safety conditioning prior to each shift.

\section{Measurement model}
At each actuation event, cells are filtered by temperature:
\begin{equation}
T_{\mathrm{lo}} \le T_i \le T_{\mathrm{hi}},
\end{equation}
where $T_{\mathrm{lo}}=\texttt{ft\_temp\_lo}$ and
$T_{\mathrm{hi}}=\texttt{ft\_temp\_hi}$. Over the filtered set, the code forms
density-weighted moments
\begin{align}
W   &= \sum_i \rho_i \Delta V_i, \\
W_z &= \sum_i \rho_i z_i \Delta V_i, \\
W_v &= \sum_i \rho_i v_{z,i}\Delta V_i.
\end{align}
The associated observables are
\begin{equation}
z_c=\frac{W_z}{W},\qquad v_c=\frac{W_v}{W}.
\end{equation}
The vertical extent of the same filtered set is also measured:
\begin{align}
z_{\min}&=\min_i z_i,\nonumber\\
z_{\max}&=\max_i z_i,\nonumber\\
z_m&=\frac{z_{\min}+z_{\max}}{2}.
\end{align}

MPI reductions are performed on $(W,W_z,W_v,z_{\min},z_{\max})$ at actuation
cadence only. Control is skipped when no filtered cells are present or when
$W\le \texttt{ft\_weight\_floor}$.

\section{Control observable and filtering}
The position signal used by the controller is selected by
\texttt{ft\_position\_signal}:
\begin{equation}
z_{\mathrm{ctrl}} =
\begin{cases}
z_c, & \texttt{centroid},\\
z_m, & \texttt{band\_midpoint},\\
(1-b)z_c + b z_m, & \texttt{blend},
\end{cases}
\end{equation}
with $b=\texttt{ft\_position\_blend}\in[0,1]$.

Actuation is attempted every \texttt{ft\_apply\_every} cycles after
\texttt{t\_frame\_tracking\_start}. Let
$\Delta t_k=t_k-t_{k-1}$ be the elapsed time since the previous control event.
An exponential moving average (EMA) is applied:
\begin{align}
\alpha_k &= 1-\exp(-\Delta t_k/\tau_{\mathrm{avg}}),\\
z_f &\leftarrow z_f + \alpha_k(z_{\mathrm{ctrl}}-z_f),\\
v_f &\leftarrow v_f + \alpha_k(v_c-v_f),
\end{align}
where $\tau_{\mathrm{avg}}=\texttt{ft\_tau\_avg}$.

Implementation detail: the first valid event initializes $(z_f,v_f)$ and
returns without actuation, preventing a startup impulse from uninitialized
filter state.

\section{Feedback law}
Define tracking error relative to the target
$z_\star=\texttt{z\_interface}$:
\begin{equation}
e_z = z_f-z_\star.
\end{equation}

The preconditioned command uses proportional, derivative, and optional integral
terms:
\begin{equation}
v_{\mathrm{pre}} = v_P + v_D + v_I.
\end{equation}

Proportional term (\texttt{position} and \texttt{pd} modes):
\begin{equation}
v_P = -\frac{e_z}{\tau_{\mathrm{relax}}},
\qquad
\tau_{\mathrm{relax}}=\texttt{ft\_tau\_relax}.
\end{equation}

Derivative/velocity term:
\begin{equation}
v_D =
\begin{cases}
-v_f, & \texttt{velocity},\\
0, & \texttt{position},\\
-\left(1-e^{-\Delta t_k/\tau_{\mathrm{vel}}}\right)v_f, & \texttt{pd},
\end{cases}
\end{equation}
with $\tau_{\mathrm{vel}}=\texttt{ft\_tau\_vel}$.

Optional integral bias-cancellation is enabled only when
\texttt{ft\_tau\_int}$>0$ and \texttt{ft\_int\_max\_abs}$>0$:
\begin{align}
v_I &\leftarrow v_I \exp(-\Delta t_k/\tau_{\mathrm{leak}}),\\
v_I &\leftarrow v_I - \frac{\Delta t_k}{\tau_{\mathrm{int}}}e_z
\quad \text{(if accumulation is allowed)},\\
|v_I| &\le v_{I,\max},
\end{align}
where
$\tau_{\mathrm{int}}=\texttt{ft\_tau\_int}$,
$\tau_{\mathrm{leak}}=\texttt{ft\_int\_leak\_tau}$, and
$v_{I,\max}=\texttt{ft\_int\_max\_abs}$.
If \texttt{ft\_int\_unsat\_only=true}, accumulation is skipped when the
pre-accumulation command magnitude already reaches the absolute boost cap,
providing anti-windup behavior.

\section{Safety conditioning and final command}
The safety sequence is applied in fixed order:
\begin{enumerate}
  \item Absolute cap:
  \begin{equation}
  |v|\le v_{\max},\qquad v_{\max}=\texttt{ft\_max\_abs\_boost}.
  \end{equation}
  \item Optional predicted-position clamp using
  \begin{equation}
  z_{\mathrm{pred}}=z_f + v\Delta t_k
  \end{equation}
  and bounds
  $[\texttt{ft\_lowzlim},\texttt{ft\_uppzlim}]$ if enabled.
  \item Slew-rate cap:
  \begin{equation}
  |v-v_{\mathrm{prev}}|\le \Delta v_{\max},
  \end{equation}
  where $\Delta v_{\max}=\texttt{ft\_max\_boost\_change}$.
  \item Re-apply absolute cap as a final guarantee.
\end{enumerate}
The resulting command is the applied shift $\Delta v_z$.

\section{Galilean state update}
For each cell with density $\rho$, the shift is applied uniformly:
\begin{align}
v_z' &= v_z + \Delta v_z,\\
m_z' &= m_z + \rho\Delta v_z.
\end{align}
For ideal EOS runs, total energy is updated conservatively:
\begin{equation}
E' = E + m_{z,\mathrm{old}}\Delta v_z + \frac{1}{2}\rho(\Delta v_z)^2.
\end{equation}
This preserves consistency with a frame-velocity translation rather than a
source-term acceleration model.

\section{Runtime interface}
The implementation is \emph{modern-only}. Frame tracking is configured with
\texttt{use\_frame\_tracking}, \texttt{t\_frame\_tracking\_start},
\texttt{z\_interface}, and the \texttt{ft\_*} parameters. Legacy keys are
explicitly rejected at startup with a migration error to prevent mixed-schema
configurations.

\begin{table}[t]
\centering
\small
\caption{Core control parameters.}
\begin{tabular}{@{}p{0.43\columnwidth}p{0.50\columnwidth}@{}}
\toprule
Parameter & Role \\
\midrule
\texttt{ft\_apply\_every} & Actuation cadence in cycles \\
\texttt{ft\_mode} & \texttt{velocity}, \texttt{position}, or \texttt{pd} \\
\texttt{ft\_tau\_avg} & EMA timescale for measured signals \\
\texttt{ft\_tau\_relax} & Proportional timescale \\
\texttt{ft\_tau\_vel} & Derivative timescale \\
\texttt{ft\_tau\_int} & Integral timescale (\(\le 0\) disables) \\
\texttt{ft\_temp\_lo/hi} & Temperature band for measurement \\
\texttt{ft\_weight\_floor} & Minimum tracked mass to actuate \\
\texttt{ft\_max\_abs\_boost} & Absolute command cap \\
\texttt{ft\_max\_boost\_change} & Slew-rate cap \\
\texttt{ft\_position\_signal} & Position-signal selector \\
\texttt{ft\_position\_blend} & Blend weight in \texttt{blend} mode \\
\bottomrule
\end{tabular}
\end{table}

\section{Diagnostics and interpretation}
At diagnostic cadence, rank 0 logs:
\begin{itemize}
  \item measured and filtered observables:
  \texttt{W\_global}, \texttt{mean\_z\_ft}, \texttt{mean\_velz\_ft},
  \texttt{z\_filt}, \texttt{vz\_filt};
  \item control decomposition:
  \texttt{v\_p}, \texttt{v\_d}, \texttt{v\_i};
  \item command outputs:
  \texttt{vel\_z\_cmd\_pre}, \texttt{vel\_z\_cmd}, \texttt{vel\_z\_shift}.
\end{itemize}
Together these fields identify lag, bias, saturation, and
observable-selection mismatch without additional instrumentation.

\section{Summary}
The TRML frame tracker is a discrete-time, globally measured Galilean controller
with explicit smoothing, optional integral bias cancellation, and ordered safety
limits. Its architecture is tuned for turbulent multiphase flows where raw
interface measurements are noisy and controller robustness must coexist with
conservative fluid updates.

\end{document}
