\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{hyperref}

\title{AthenaK CGL--Landau-Fluid STS Validation Note}
\author{}
\date{April 24, 2026}

\begin{document}
\maketitle

\section{Purpose}

This note documents the AthenaK CGL--Landau-fluid (CGL--LF) heat-flux update
implemented on branch \texttt{CGL-STS-LF}.  The goal is not to reproduce the full
nonlinear turbulence campaign of Squire et al. (2023), but to show that the
implemented method follows the same finite-volume variable strategy and can
reproduce the method-level checks, including a one-dimensional oblique
linear-wave setup using the Figure 17 background state.

\section{Conserved Variables}

The hyperbolic CGL system evolves total energy and the conserved anisotropy
variable
\begin{equation}
  A = \rho \log\left[
    \frac{p_\perp}{p_\parallel}
    \frac{\rho^2}{|B|^3}
  \right].
\end{equation}
The legacy array slot is still named \texttt{IMU}, with \texttt{IAN} used as an
alias in CGL code.  The Riemann solve remains in \(A\), which avoids the
small-\(|B|\) pathologies associated with evolving magnetic moment directly.

The LF STS operator temporarily changes only the interpretation of the
\texttt{IAN} slot:
\begin{equation}
  A \rightarrow \mu_B = \frac{p_\perp}{|B|}.
\end{equation}
The standard conserved-variable STS update then advances \((E,\mu_B)\) with the
LF flux divergence.  Immediately after the RKL2 stage, the EOS converts
\(\mu_B \rightarrow A\), and the normal CGL conserved-to-primitive recovery
restores \((p_\parallel,p_\perp)\).

\section{Landau-Fluid Closure}

The implemented closure uses the Snyder--Hammett--Dorland 3+1 heat fluxes,
\begin{align}
  q_\parallel &=
    -\chi_\parallel \rho \nabla_\parallel\left(\frac{p_\parallel}{\rho}\right),
  \\
  q_\perp &=
    -\chi_\perp \left[
      \rho \nabla_\parallel\left(\frac{p_\perp}{\rho}\right)
      - p_\perp\left(1-\frac{p_\perp}{p_\parallel}\right)
        \frac{\nabla_\parallel |B|}{|B|}
    \right],
\end{align}
with
\begin{equation}
  \chi_\perp = \sqrt{\frac{2}{\pi}}\frac{c_\parallel}{|k_\parallel|},
  \qquad
  \chi_\parallel = \sqrt{\frac{8}{\pi}}\frac{c_\parallel}{|k_\parallel|}.
\end{equation}
Here \texttt{lf\_k\_parallel} is the closure wavenumber \(|k_\parallel|\), not a
conductivity.  In local mode, \(c_\parallel=\sqrt{p_\parallel/\rho}\) is computed
from the live face state.  In background mode, only this prefactor is frozen to
\texttt{lf\_c\_parallel0}; all gradients and thermodynamic variables remain live.

The face fluxes coupled into STS are
\begin{equation}
  F_E^i = \hat{b}^i \left(q_\perp + \frac{1}{2}q_\parallel\right),
  \qquad
  F_{\mu_B}^i = \hat{b}^i \frac{q_\perp}{|B|}.
\end{equation}

\section{Heat-Flux Cap}

Following Squire et al. (2023), equations 3.1--3.2, the unlimited heat flux
\(q_L\) is capped as
\begin{equation}
  q = \frac{q_L q_{\max}}{q_{\max} + |q_L|}.
\end{equation}
The implemented caps are
\begin{equation}
  q_{\perp,\max} = \sqrt{\frac{2}{\pi}} c_\parallel p_\perp,
  \qquad
  q_{\parallel,\max} = \sqrt{\frac{8}{\pi}} c_\parallel p_\parallel.
\end{equation}
Collision-frequency-dependent suppression of the coefficients is intentionally
left for a later pass.

\section{Validation Problems}

\begin{table}[h]
\centering
\small
\begin{tabular}{p{0.34\textwidth}p{0.32\textwidth}p{0.24\textwidth}}
\toprule
Input & Purpose & Current result \\
\midrule
\texttt{cgl\_transform\_check} & \(A \leftrightarrow \mu_B\) and primitive recovery
  & pass \\
\texttt{cgl\_lf\_quant\_parallel} & sinusoidal \(T_\parallel\) decay
  & \(8.1854{\times}10^{-5}\) vs \(8.1830{\times}10^{-5}\) \\
\texttt{cgl\_lf\_quant\_perp} & sinusoidal \(T_\perp\) decay
  & \(9.0469{\times}10^{-5}\) vs \(9.0460{\times}10^{-5}\) \\
\texttt{cgl\_lf\_quant\_grad\_b} & isolated \(\nabla_\parallel |B|\) term
  & RMS relative error \(1.01{\times}10^{-2}\) \\
\texttt{cgl\_lf\_flux\_limiter} & equation 3.2 heat-flux cap
  & RMS relative error \(1.12{\times}10^{-2}\) \\
\texttt{cgl\_lf\_limiter\_mirror} & mirror-biased admissibility stress
  & pass \\
\texttt{cgl\_lf\_limiter\_firehose} & firehose-biased admissibility stress
  & pass \\
\texttt{cgl\_lf\_field\_wave} & field-aligned LF wave reference
  & pass \\
\texttt{cgl\_lf\_paper\_oblique\_wave} & Figure 17 background, LF enabled
  & pass \\
\texttt{cgl\_pure\_paper\_oblique\_wave} & Figure 17 background, no LF
  & pass \\
\bottomrule
\end{tabular}
\end{table}

\section{Figure 17-Style Linear Wave}

The paper-style inputs use the Figure 17 background
\begin{equation}
  \rho_0 = 1,\quad p_{\parallel 0}=p_{\perp 0}=5,\quad
  u_0=0,\quad B_0=(1,\sqrt{2},0.5),\quad k=2\pi \hat{x},
  \quad kL=|k|.
\end{equation}
The test initializes a small transverse velocity perturbation and compares the
final complex Fourier amplitudes of \(u_y\), \(B_y\), \(p_\parallel\), and
\(p_\perp\) against an offline linearized CGL--LF reference integrated with the
same background state and closure coefficients.  Two inputs are provided: one
with the LF heat flux enabled and one with pure CGL.  This checks the same
method class as the Appendix A.2/Figure 17 linear-wave convergence problem,
while leaving exact eigenmode initialization and full convergence scans for a
follow-up once Steve/Jono provide canonical eigenvectors or damping rates.

\section{Build and Runtime Status}

The CPU quantitative build completed successfully.  The only build warnings
observed were pre-existing variable-length-array warnings in \texttt{mesh.cpp}.
No CGL-owned files were deleted.

\begin{thebibliography}{9}
\bibitem{Squire2023}
J. Squire et al.,
\emph{Pressure anisotropy and viscous heating in weakly collisional plasma
turbulence},
\href{https://arxiv.org/abs/2303.00468}{arXiv:2303.00468}, 2023.
\end{thebibliography}

\end{document}
